Hydrogen Epoch of Reionization Array (HERA)

### 2016-04-20 manuscript.tex
incorporate (most of) Jonnies comments

The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR). During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.  This paper will provide an overview of the project and describe the design of the system.

\section{Introduction}
The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
other probes, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of a new understanding of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).

Because of these key science questions that may be probed by highly redshifted hydrogen, there has been a resurgence in low frequency arrays  -- most with a primary
objective of measuring the EOR.  These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in Sec. \ref{sec:delayapproach}).
At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
Examples include improving foreground subtraction to expand the information content 
in power spectral and imaging measurements and
extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).

@@

The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more 
sensitivity than current instruments and provide a significant chance at detecting the EoR power spectrum signal.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER \citep{ali_et_al2015}, which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).  The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).  The white dashed horizontal lines are the HERA EoR band and the blue are the extended frequency goal.  Determining where the structure in the background really falls is the
goal of HERA.} 
	\label{fig:cosmos}
\end{figure}

Exploring the interplay of galaxies and large-scale structure during the EOR
requires complementary observational approaches. Measurements of the Cosmic Microwave Background 
(CMB; the escaping radiation permeating the universe after the universe became
transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.  Thomson
scattering of CMB photons by the ionized particles constrains the integrated column of 
ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the power spectrum of cosmic structures.
But even with these measurements, the detailed evolution of the IGM 
is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,zahn_etal2012,mesinger_et_al2012}.
Lyman-$\alpha$ absorption features in quasar and $\gamma$-ray burst spectra give 
ionization constraints at the tail end of reionization 
($z\la7$, \citealt{fan_et_al2006, mcgreer_et_al2015}), but these features 
saturate at low neutral fractions
($x_{\rm HI} \la 10^{-4}$), where $x_{\rm HI}$ is the ratio of neutral hydrogen to the total baryonic density.

@@

\begin{figure}[h!]
\centering
    \includegraphics[width=0.49\textwidth,clip]{plots/sensitivity_wideband.pdf}
  \caption{ 1$\sigma$ thermal noise errors on $\Delta^2(k)$, the 21\,cm power spectrum, at $k\!=\!0.2$\,$h$\,Mpc$^{-1}$ (the dominant error at that $k$)
with 1080 hours of integration (black)
compared with various heating and reionization models (colored).  Note that only predicted results using the delay-spectrum approach are shown.}
	\label{fig:Sensitivities}
\end{figure}

HERA builds on the advances of first-generation
21\,cm EoR experiments led by HERA team members, particularly 
the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}),
the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).
Recent measurements
have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, parsons_et_al2015, ali_et_al2015,pober_et_al2015}.
However, current experiments cannot expect more than marginal detections of the EoR signal. 
Figure~\ref{fig:Sensitivities} compares telescope sensitivities as a function of redshift to models of 
the evolving, spherically averaged 21\,cm EoR power spectrum, characterized by the dimensionless power spectrum parameter $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$.
The expected performance of HERA relative to current and planned telescopes to detect the peak of reionization (as well as the total collecting area is shown in 
Table~\ref{tab:signif}.

@@

\begin{SCtable}
\caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
\small
 \centering
 \begin{tabular}{c||r||r|r} 
%\begin{deluxetable}{c||r||r|r}
%\tabletypesize{\small}
%\tablecaption{\small
\hline
%\startdata
Instrument & \shortstack{Collecting \\ Area (m$^2$)} & \shortstack{Foreground \\Avoidance} & \shortstack{Foreground \\Modeling} \\
\hline
PAPER & 1,188 & 0.77$\sigma$ & 3.04$\sigma$ \\
MWA & 3,584 & 0.31$\sigma$ & 1.63$\sigma$ \\
LOFAR NL Core & 35,762 & 0.38$\sigma$ & 5.36$\sigma$ \\
%\textbf{HERA-127} & \textbf{19,500} & \textbf{10.88} & \textbf{35.65} \\
\textbf{HERA-350} & \textbf{53,878} & \textbf{23.34$\sigma$} & \textbf{90.97}$\boldsymbol{\sigma}$ \\
SKA1 Low Core & 416,595 & 13.4$\sigma$ & 109.90$\sigma$
%\enddata
\end{tabular}
%\Caption{-0.1in}{0.99}{-0.1in}
\hspace{-0.1in}
\label{tab:signif}
\end{SCtable}

@@

HERA's 21\,cm measurements can be used in conjunction with our semi-analytic models to constrain the ionization history. 
The red band in Fig.~\ref{fig:IonHist} shows the forecasted 95\% confidence region derived from HERA data after marginalizing over astrophysical and cosmological parameters.
Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Sec. \ref{sec:freqs}).
Based on the excursion-set formalism of
\citet{furlanetto_et_al2004} and the 21cmFAST code \citep{mesinger_et_al2011},
this code models the astrophysics of
reionization with three free parameters (see Figure~\ref{fig:paramConstraints} for details). 
While the existing experiment with the most collecting area, 
the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V, yatawatta_et_al2013}),
provides some ability to constrain these parameters,
HERA's constraints are significantly more precise and are comparable
to what could be achieved with the SKA. 
Additionally, HERA's constraints enable principal component parameterizations of the
sky-averaged $21\,\textrm{cm}$ signal measurements pursued by experiments such as EDGES, increasing their signal-to-noise and thus their science return \citep{liu_parsons2015}.

@@

\section{Measuring the EOR}
\label{sec:eormeas}
Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.  The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire universe back almost to the period of recombination.  The bandwidth
therefore equates to a cosmic distance along the line-of-site of the telescope, with the frequency determining the cosmic age.  An observation of an area of sky over a given bandwidth is therefore providing an average
over a cosmic volume.

@@

The past three years have seen deep EoR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
The inherent challenge is to simultaneously meet stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).

A promising approach to make the first detection of the EoR power spectrum 
is a foreground mitigation strategy based largely
on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  This is the approach
pioneered by PAPER.
By reducing the need
for foreground modeling and subtraction, this approach allowed PAPER to switch to a grid-based antenna layout that
enhanced sensitivity toward fewer Fourier modes and facilitated calibration
based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{liu_et_al2010,zheng_et_al2014}.
Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
to the foreground problem.

Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
leave it with limited diagnostic capability, particularly for direction-dependent systematics
such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013, nunhokee_etal_2016}.
While concern over such effects has decreased markedly since discovering that
intrinsic point source polarization is lower than previously thought
\citep{asad_et_al2015} and that variable Faraday rotation through the
ionosphere averages down the polarized signal over long observing seasons
\Mycitep{moore_et_al2016,aguirre_etal_2016}, direction-dependent beam effects remain an area of interest.
The MWA's image-based calibration and foreground subtraction strategy provides complementary
capabilities (Sec. \ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
can increase sensitivity by recovering more modes of the cosmological signal (see
Table~\ref{tab:signif}) and help address systematic errors rooted
in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
the proven approaches of redundant calibration and delay filtering, while simultaneously
increasing the extent and density of $uv$ sampling for high-fidelity imaging.

@@

The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response.   The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.  Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.  For HERA, wavenumbers are dominated by the bandwidth, not the baseline.

Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmos by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.  The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    
Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.

\begin{figure}[h!]
\centerline{
\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
}
\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.  The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).  See text for discussion.  The full redshift range is the HERA extended goal.  The initial HERA delay-spectrum power spectrum goal is the light green square region in the upper left.}
\label{fig:kperf}
\end{figure}

In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.  Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
\begin{equation}
\kpar \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
\end{equation}
where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included(see Fig. \ref{fig:wedge}).

@@

Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).  In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.  Also note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.

@@

\subsection{Delay-Spectrum Approach}
\label{sec:delayapproach}
Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.  The delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.  Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.

>>>>

The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment that uses the unique properties of the 21\,cm line from neutral hydrogen to probe the Epoch of Reionization (EOR). During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.  This paper will provide an overview of the project and describe the design of the system, moving from a science overview to requirements, design and status.

\section{Introduction}
The Hydrogen Epoch of Reionization Array (HERA; http://reionization.org) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
other probes, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of a new understanding of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).

Because of these key science questions that may be probed by highly-redshifted hydrogen emission, there has been a resurgence in low frequency arrays  -- most with a primary
objective of measuring the EOR.  These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in \S\ref{sec:delayapproach}).
At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
Examples include improving foreground subtraction to expand the information content 
in power spectral and imaging measurements and
extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).

@@

The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide a factor of about 5 more 
sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
%\citep{ali_et_al2015}
which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.  These estimates will be quantified and shown in this paper.

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).  The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).  The white dashed horizontal lines are the HERA EOR band and the blue are the extended frequency goal.  Determining where the structure in the background really falls is the
goal of HERA.} 
	\label{fig:cosmos}
\end{figure}

Exploring the interplay of galaxies and large-scale structure during the EOR
requires complementary observational approaches. Measurements of the Cosmic Microwave Background 
(CMB; the photons permeating the universe after becoming
transparent to its own radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.  Thomson
scattering of CMB photons by the ionized particles constrains the integrated column of 
ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the duration of the ``patchy'' phase of cosmic structures.
But even with these measurements, the detailed evolution of the IGM 
is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,zahn_etal2012,mesinger_et_al2012}.
Lyman-$\alpha$ absorption features in quasar and $\gamma$-ray burst spectra give 
ionization constraints at the tail end of reionization 
($z\la7$, \citealt{fan_et_al2006, mcgreer_et_al2015}), but these features 
saturate at low neutral fractions
($x_{\rm HI} \ga 10^{-4}$), where $x_{\rm HI}$ is the fraction of hydrogen in its neutral state.

@@

\begin{figure}[h!]
\centering
    \includegraphics[width=0.49\textwidth,clip]{plots/sensitivity_wideband.pdf}
  \caption{ 1$\sigma$ thermal noise errors on $\Delta^2(k)$, the 21\,cm power spectrum, at $k\!=\!0.2$\,$h$\,Mpc$^{-1}$ (the dominant error at that $k$)
with 1080 hours of integration (black)
compared with various heating and reionization models (colored).  Sensitivity analysis is per footnote in Table \ref{sec:signif}}
	\label{fig:Sensitivities}
\end{figure}

HERA builds on the advances of first-generation
21\,cm EoR experiments led by HERA team members, particularly 
the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}),
the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).
Recent measurements
have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, parsons_et_al2015, ali_et_al2015,pober_et_al2015}.
However, current experiments cannot expect more than marginal detections of the EoR signal. 
Figure~\ref{fig:Sensitivities} compares telescope sensitivities as a function of redshift to models of 
the evolving, spherically averaged 21\,cm EoR power spectrum, characterized by the dimensionless power spectrum parameter $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$.
The expected performance of HERA relative to current and planned telescopes to detect the peak of reionization (as well as the total collecting area is shown in 
Table~\ref{tab:signif}).

@@

\begin{table}
\caption{\small Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8^*$.
}
\small
 \centering
 \begin{tabular}{c||r||r|r} 
%\begin{deluxetable}{c||r||r|r}
%\tabletypesize{\small}
%\tablecaption{\small
\hline
%\startdata
Instrument & \shortstack{Collecting \\ Area (m$^2$)} & \shortstack{Foreground \\Avoidance} & \shortstack{Foreground \\Modeling} \\
\hline
PAPER & 1,188 & 0.77$\sigma$ & 3.04$\sigma$ \\
MWA & 3,584 & 0.31$\sigma$ & 1.63$\sigma$ \\
LOFAR NL Core & 35,762 & 0.38$\sigma$ & 5.36$\sigma$ \\
%\textbf{HERA-127} & \textbf{19,500} & \textbf{10.88} & \textbf{35.65} \\
\textbf{HERA-350} & \textbf{53,878} & \textbf{23.34$\sigma$} & \textbf{90.97}$\boldsymbol{\sigma}$ \\
SKA1 Low Core & 416,595 & 13.4$\sigma$ & 109.90$\sigma$
%\enddata
\end{tabular}
\caption*{%\tiny{\justify
$^*$Calculations done via \url{21cmSense} \url{www.github.com/} \citep{2013AJ....145...65P,2014ApJ...782...66P}
Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows significantly more $k$ modes of the cosmological signal to be recovered.}
\label{tab:signif}
\end{table}

@@

HERA's 21\,cm measurements can be used in conjunction with our semi-analytic models to constrain the ionization history. 
The red band in Fig.~\ref{fig:IonHist} shows the forecasted 95\% confidence region derived from HERA data after marginalizing over astrophysical and cosmological parameters.
Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see \S\ref{sec:freqs}).
Based on the excursion-set formalism of
\citet{furlanetto_et_al2004} and the 21cmFAST code \citep{mesinger_et_al2011},
this code models the astrophysics of
reionization with three free parameters (see Figure~\ref{fig:paramConstraints} for details). 
While the existing experiment with the most collecting area, 
the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V, yatawatta_et_al2013}),
provides some ability to constrain these parameters,
HERA's constraints are significantly more precise and are comparable
to what could be achieved with the SKA. 
Additionally, HERA's constraints enable principal component parameterizations of the
sky-averaged $21\,\textrm{cm}$ signal measurements pursued by experiments such as EDGES, increasing their signal-to-noise and thus their science return \citep{liu_parsons2015}.

@@

\section{Measuring the EOR}
\label{sec:eormeas}
Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the redshift of spectral lines.  The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire universe back almost to the period of recombination.  The bandwidth
therefore equates to a cosmic distance along the line-of-site of the telescope, with the frequency determining the cosmic age.  An observation of an area of sky over a given bandwidth is therefore providing an average
over a cosmic volume.

@@

The past three years have seen deep EOR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
The inherent challenge is to simultaneously meet stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}), handling foreground contamination
({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).

A promising approach to make the first detection of the EoR power spectrum 
is a foreground mitigation strategy based largely
on identifying and filtering out a region of parameter space where the strong foregrounds are largely confined (\S\ref{sec:delayapproach}).  This is the approach
pioneered by PAPER.
By reducing the need
for foreground modeling and subtraction, this approach allowed PAPER to switch to a grid-based antenna layout that
enhanced sensitivity toward fewer Fourier modes and facilitated calibration
based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{liu_et_al2010,zheng_et_al2014}.
Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
to the foreground problem.

Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
leave it with limited diagnostic capability, particularly for direction-dependent systematics
such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013, nunhokee_etal_2016}.
While concern over such effects has decreased markedly since discovering that
intrinsic point source polarization is lower than previously thought
\citep{asad_et_al2015} and that variable Faraday rotation through the
ionosphere averages down the polarized signal over long observing seasons
\Mycitep{moore_et_al2016,aguirre_etal_2016}, direction-dependent beam effects remain an area of interest.
The MWA's image-based calibration and foreground subtraction strategy provides complementary
capabilities (\S\ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
can increase sensitivity by recovering more modes of the cosmological signal (see
Table~\ref{tab:signif}) and help address systematic errors rooted
in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
the proven approaches of redundant calibration and delay filtering, while simultaneously
increasing the extent and density of $uv$ sampling for high-fidelity imaging.

@@

The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response \footnote{Note that this equivalency is an approximation, which is very good for the short baselines and small bandwidths generally used here.  For a more complete discussion see \cite{liu_et_al2014a,liu_et_al2014b}}.   
The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.  Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.  For HERA, wavenumbers are dominated by the bandwidth, not the baseline.

Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmology by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.  The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    
Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.

\begin{figure}[h!]
\centerline{
\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
}
\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.  The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-sight redshift ranges for that bandwidth (too small to see at 1 MHz).  See text for discussion.  The full redshift range is the HERA extended goal.  The initial HERA delay-spectrum power spectrum goal is the light green square region in the upper left.}
\label{fig:kperf}
\end{figure}

In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr, \kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.  Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
\begin{equation}
\kpar \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
\end{equation}
where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included (see Fig. \ref{fig:wedge}).

@@

Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).  In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.  Also note that at the redshifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.

@@

\subsection{Delay-Spectrum Approach}
\label{sec:delayapproach}
Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.  The delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.  Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.  For more details, see \cite{parsons_et_al2012b}.

We can essentially relate the sky power spectrum $P(\kvec)$ to the Fourier transform along the frequency axis (the delay-transform) of an interferometer baseline visibility, denoted ${\tilde V}_b(\tau)$:
\begin{equation}
P(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2_b(\tau)}{\Omega_b B/ \lambda^4} \right]
\label{eq:Pk_baseline}
\end{equation}
where $X$ and $Y$ were introduced above, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.  The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.  
Rather than $P(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}P(\kvec)$.
